A general and efficient representation of Ancestral Recombination Graphs

### 2023-08-30 paper.tex
Refine

\section*{Introduction}
% Inferring the genetic relationships between sampled genomes in the form of a
% phylogenetic tree is a necessary prerequisite for many analyses of
% species~\citep{rannala2003genetics}, non-recombining sections of
% DNA~\citep{cann1987mitochondrial,underhill2001annalsofhumangenetics}
% and viruses~\citep{grenfell2004science}.
% A phylogenetic tree captures information about the genealogy
% of the sample in an elegant and concise
% way, by postulating a set of ancestors common to the samples
% together with paths of inheritance between them. A rich literature exists
% on analysing the mathematical properties of these
% trees~\citep{steel2016phylogeny}, and numerous
% methods exist to infer them~\citep{felsenstein2004inferring}.
% However, the genetic genealogy of a recombining organism
% cannot be captured by a single tree: instead,
% the inheritance of each section of the genome may be described
% by a (slightly) different local tree~\citep{hudson1983properties}.
% Inferring complex
% genealogies of this nature
% presents profound technical difficulties, meaning genealogy-based approaches have
% not been widely used in population and statistical genetics.
% Until recently, analysis of
% population genetic data has therefore primarily focussed on summaries derived
% from the local trees~\citep{tajima1983evolutionary,tavare1984line}
% rather than using the full underlying genealogy itself. Such
% derived measures include allele frequencies and site frequency
% spectra~\citep{achaz2009frequency,ralph2020efficiently},
% patterns of linkage disequilibrium~\citep{mcvean2002genealogical}, and
% principal components~\citep{mcvean2009genealogical}.
Estimating the detailed genetic genealogy of a set of sequences
under the influence of recombination,
usually known as an Ancestral Recombination Graph (ARG), is a long-standing
goal in genetics.
Recent breakthroughs
in large-scale inference
methods~\citep{rasmussen2014genome,kelleher2019inferring,speidel2019method,
schaefer2021ancestral,wohns2022unified,zhang2023biobank,zhan2023towards}
and data representation~\citep{kelleher2016efficient,kelleher2018efficient}
have raised the realistic prospect of ARG-based analysis becoming a standard part
of the population and statistical genetics toolkit~\citep{hejase2020summary}.
Applications using these inferred ancestries as inputs have begun to
appear~\citep{osmond2021estimating,fan2022genealogical,hejase2022deep,zhang2023biobank,
% TODO update when natgen paper comes out
nowbandegani2022extremely}
and many more are sure to
follow~\citep{harris2019database,harris2023using}.

@@

Prospective ARGs occur in individual-based forwards-time
simulations~\citep{kelleher2018efficient,haller2018tree},
where the genetic inheritance between genomes is recorded
exhaustively and periodically
``simplified'' to retain only
the ancestry relevant to the current population.
\citet{kelleher2018efficient} refer to the structure
recording all such forwards-in-time genetic relationships
as an ``embellished pedigree`` (and suggest the term ``nedigree''
as a shorthard).
Describing this instead as a ``prospective ARG'', emphasising
the use of the same data structures and the direct connection with
the usual retrospective interpretation of an ARG, may be
more useful.
The ``simplify'' algorithm which converts a prospective
gARG into a retrospective one is a powerful tool,
with several other applications
as we discuss in the following sections.

\section*{Converting an eARG to a gARG}
Any eARG can be converted to a gARG without loss of information,
and the conversion process is very similar to
simplifying a prospective gARG with respect to a set of samples.
The process involves two steps, and is illustrated in
Fig~\ref{fig-ancestry-resolution}.
The first step is straightforward
(Figure~\ref{fig-ancestry-resolution}B).
We duplicate the nodes and
edges from the input eARG, and the add inheritance annotations
to the output gARG's edges. If the node is a common ancestor node,
we annotate the single outbound edge with the interval $(0,L]$,
for a genome of length $L$.
If the node is a recombination event with a breakpoint $x$,
we annotate the outbound edges with the intervals $(0, x]$
and $(x, L]$.
These inheritance annotations are clearly in one-to-one
correspondance with the information in the input eARG,
but they are a superset of the
``ancestral material''~\citep{wiuf1999ancestry,wiuf1999recombination}
of the sample.

\begin{figure}
\centering
% TODO update the source of this figure to make less tall. It's taking
% up far too much vspace at in native coords at the moment.
\includegraphics[width=0.75\textwidth]{illustrations/ancestry-resolution}
\caption{\label{fig-ancestry-resolution}
Converting the \citet[][Fig.~1]{wiuf1999recombination} example graph
to a gARG. (A) The original eARG; square nodes represent events, with
each recombination (red) showing the corresponding breakpoint.
(B) The corresponding gARG with breakpoints directly converted to
edges annotated with inheritance intervals.
(C) The sample-resolved gARG resulting from simplifying with respect
to the samples, $a$, $b$, and $c$. Dashed lines show edges that are
no longer present (in practice, nodes $g$ and $q$ would also be removed).
Coalescence with respect to the sample is indicated by shaded bars, as
in Fig~\ref{fig-arg-in-pedigree}A.
Line thickness is proportional to the genomic span of each edge.
Nodes representing recombination events have been retained
for clarity, but could be removed by simplification if
desired.
}
\end{figure}

Ancestral material is a critical concept in a retrospective ARG,
and refers to the genomic intervals directly ancestral to the samples
scattered across lineages (the edges of the graph)
at some time in the past. At recombination
events, ancestral material is split between the ancestral lineages.
As genome segments coalesce in a common
ancestor the total amount ancestral material is reduced, until
eventually as we traverse pastwards through the graph, all
samples reach a most recent common ancestor at all locations
along the genome. The process of transmitting ancestral material
along the edges of an ARG is directly analogous to Hudson's
simulation algorithm for the coalescent with
recombination~\citep{hudson1983testing,kelleher2016efficient}.

Fig~\ref{fig-ancestry-resolution}C shows the graph that we obtain
by resolving the ancestral material transmitted along each edge
(using the ``simplify'' algorithm), and highlights some important
differences between the eARG and gARG encodings.
Firstly, we can see that some nodes and edges are removed entirely
from the graph. The ``grand MRCA'' \noderef{q} is omitted from the
fully resolved gARG because all segments of the genome have
fully coalesced before it is reached. Likewise, the edge
between \noderef{g} and \noderef{j} is omitted because the recombination
event at position $5$ (represented by node \noderef{g})
fell in non-ancestral material.
More generally, we can see that the fully resolved
gARG encoding of Fig~\ref{fig-ancestry-resolution}C
allows for ``local'' inspection
of an ARG in a way that is not possible by storing
common ancestor and recombination events in an eARG. Because
the ancestral material is stored with each edge, the
cumulative effects of events over time can be reasoned
about at once. Many computations
that we wish to perform on an ARG will require resolving
the genetic material with respect to a sample.
% this is weak, but would be good to emphasise this point.
% you just end up running simplify lots of times, in your
% downstream algs
The gARG encoding
allows us to do this once and to store the result, whereas
the eARG encoding requires us to repeat the process on
each traversal.

It is important to note here that the \citet{wiuf1999recombination} eARG
example used here is something of a straw man, in that it is highly unlikely
that an inference or simulation program would produce a retrospective
ARG that contains
nodes or edges that are not ancestral to the sample. Nonetheless, it
is an instructive example from the literature which highlights several
important properties of ARGs, and the general point about
the need to resolve ancestral material ``on the fly'' for eARG traversals
holds.

@@

\section*{Computational efficiency and data interchange}
ARG inference methods have been successfully applied to vast datasets,
including 500,000 genotyped humans~\citep{kelleher2019inferring,zhang2023biobank}
from UK Biobank~\citep{bycroft2018genome}
and over a million SARS-Cov-2 whole genomes~\citep{zhan2023towards}.
Even larger datasets are
available~\citep[e.g.][]{halldorsson2022sequences} % TODO more
and it seems inevitable that ARGs will be inferred for them.
The efficiency with which we can store and analyse these inferred
ARGs is therefore of great importantance, and a key factor
in the ultimate success of downstream ARG-based methods.
Another key factor in how important ARG-based methods become
across the wide range of predicted application
% TODO more?
areas~\citep{rasmussen2014genome,harris2019database,
hejase2020summary,schaefer2021ancestral,
fan2022genealogical,wohns2022unified,harris2023using,
nowbandegani2022extremely}
is how well inference methods and downstream analysis applications
interoperate. Without a well-defined and efficient shared format
for ARG data interchange,
users will suffer from problems of converting data from one programs
output to another, an unfortunate hallmark of population genetics
software for many years~\citep{excoffier2006computer}.
While current methods tend to tightly couple downstream analysis
of the inferred ARG with the inference
itself within the same software package
% FIXME-  more explicit?
(e.g.\ Relate, ARGneedle)
this approach scales poorly in terms of software development
effort, and is ultimately not compatible with the widespread use
of ARGs for routine data analysis, and a healthy and diverse
software ecosystem.
% FIXME not sure where this goes
Earlier efforts to standardise ARG interchange
did not succeed~\citep{cardona2008extended,mcgill2013graphml}.

The gARG encoding discussed and concretely defined in this manuscript
leads to highly efficient storage and processing of ARG data,
and has already been in use for several years.
The succinct tree sequence data structure (usually known as a ``tree sequence''
for brevity, but see below for confusion around this point)
is a concrete encoding of a gARG, as discussed in this article.
It was originally developed as part of the \texttt{msprime}
simulator~\citep{kelleher2016efficient} and subsequently been
extended and applied to forward-time
simulations~\citep{kelleher2018efficient,haller2018tree},
calculation of population genetics statistics~\citep{ralph2020efficiently}.
The succinct tree sequence encoding extends the basic definition
of a gARG provided here by stipulating a
simple tabular representation and nodes and edges,
and also defining a concise and lossless representation of
sequence variation using the ``site'' and  ``mutation'' tables.
The \texttt{tskit} library~\citep{tskit2023tskit} is a liberally
licensed open source toolkit that provides a comprehensive suite
of tools for working with gARGs (encoded as a succinct tree sequence).
Based on core functionality written
in C, it provides interfaces in C, Python and Rust.
Tskit is mature software, widely used in population genetics, and
has been incorporated into several downstream
applications~\citep[e.g.,][]{haller2019slim,speidel2019method,
adrion2020community,
terasaki2021geonomics,
baumdicker2021efficient,
fan2022genealogical,korfmann2022weak,
mahmoudi2022bayesian,petr2022slendr,rasmussen2022espalier,
zhang2023biobank}.

% FIXME this needs more work and doesn't really fit into the
% narrative right now, but we do need to make this point.
The key insight that makes the succinct tree sequence encoding
an efficient substrate for defining analysis algorithms is that
it allows us to generate the local trees along the genome
in a way that allows us to reason about the \emph{differences}
between those trees.
Sequentially generating the marginal trees along the genome
is also fundamental, and is necessary whenever we need to
perform calculations that are contingent on more than just the
isolated properties of an edge. \cite{kelleher2016efficient}
showed how all trees can be sequentially generated in
constant time per tree transition in a fully simplified gARG.
Furthermore, we can easily reason about how tree topologies
change (and stay the same), leading to efficient algorithms
for computing population genetic
statistics~\citep{kelleher2016efficient,kelleher2018efficient},
implementing the Li and Stephens
model~\citep{kelleher2019inferring,wohns2022unified}
and likely many more.

In retrospect, the choice of terminology around the succinct
tree sequence data structure was unfortunate, and has led to
significant confusion. As first introduced by~\cite{kelleher2016efficient}
a ``tree sequence`` was defined as the set of ``coalescence records``
that is output by Hudson's simulation algorithm.
This was then generalised to use the
tabular encoding described above and formally described as a ``succinct
tree sequence''~\citep{kelleher2018efficient}.
The ``succinct'' prefix here is intended to connect
with the concept of succinct data structures
\citep[e.g.][]{gog2014theory},
which have near-optimal
space usage but still support efficient retrieval and computation.
The concrete data structures of the succinct tree sequence, precisely
defined and with a particular focus on computational efficiency
did not have an obvious connection to the ARG definitions
provided by Griffiths and
colleagues~\citep{griffiths1991two,ethier1990two,
griffiths1996ancestral,griffiths1997ancestral}, and with the lack
of any other precise definitions, it seemed best to avoid the
term to prevent confusion. Unfortunately, however, the opposite
has occured. In describing the output of \tsinfer\ as a ``tree
sequence''~\citep{kelleher2019inferring} many
have---not unreasonably, but incorrectly---concluded that the
output is a collection of \emph{independent}
% MORE?
% https://github.com/tskit-dev/what-is-an-arg-paper/issues/38
trees~\citep[e.g.,][]{hejase2020summary,ignatieva2021kwarg}.
% Could say more here, but maybe we'll have a note about terminology
% in the discussion

>>>>

\section*{Introduction}
% Inferring the genetic relationships between sampled genomes in the form of a
% phylogenetic tree is a necessary prerequisite for many analyses of
% species~\citep{rannala2003genetics}, non-recombining sections of
% DNA~\citep{cann1987mitochondrial,underhill2001annalsofhumangenetics}
% and viruses~\citep{grenfell2004science}.
% A phylogenetic tree captures information about the genealogy
% of the sample in an elegant and concise
% way, by postulating a set of ancestors common to the samples
% together with paths of inheritance between them. A rich literature exists
% on analysing the mathematical properties of these
% trees~\citep{steel2016phylogeny}, and numerous
% methods exist to infer them~\citep{felsenstein2004inferring}.
% However, the genetic genealogy of a recombining organism
% cannot be captured by a single tree: instead,
% the inheritance of each section of the genome may be described
% by a (slightly) different local tree~\citep{hudson1983properties}.
% Inferring complex
% genealogies of this nature
% presents profound technical difficulties, meaning genealogy-based approaches have
% not been widely used in population and statistical genetics.
% Until recently, analysis of
% population genetic data has therefore primarily focussed on summaries derived
% from the local trees~\citep{tajima1983evolutionary,tavare1984line}
% rather than using the full underlying genealogy itself. Such
% derived measures include allele frequencies and site frequency
% spectra~\citep{achaz2009frequency,ralph2020efficiently},
% patterns of linkage disequilibrium~\citep{mcvean2002genealogical}, and
% principal components~\citep{mcvean2009genealogical}.
Estimating the detailed genetic genealogy of a set of sequences
under the influence of recombination,
usually known as an Ancestral Recombination Graph (ARG), is a long-standing
goal in genetics.
Recent breakthroughs
in large-scale inference
methods~\citep{rasmussen2014genome,kelleher2019inferring,speidel2019method,
schaefer2021ancestral,wohns2022unified,zhang2023biobank,zhan2023towards}
and data representation~\citep{kelleher2016efficient,kelleher2018efficient}
have raised the realistic prospect of ARG-based analysis becoming a standard part
of the population and statistical genetics toolkit~\citep{hejase2020summary}.
Applications using these inferred ancestries as inputs have begun to
appear~\citep{osmond2021estimating,fan2022genealogical,hejase2022deep,zhang2023biobank,
nowbandegani2023extremely}
and many more are sure to
follow~\citep{harris2019database,harris2023using}.

@@

Prospective ARGs occur in individual-based forwards-time
simulations~\citep{kelleher2018efficient,haller2018tree},
where the genetic inheritance between genomes is recorded
exhaustively and periodically
``simplified'' to retain only
the ancestry relevant to the current population.
\citet{kelleher2018efficient} refer to the structure
recording all such forwards-in-time genetic relationships
as an ``embellished pedigree`` (and suggest the term ``nedigree'' as a shorthand).
Describing this instead as a ``prospective ARG'', emphasising
the use of the same data structures and the direct connection with
the usual retrospective interpretation of an ARG, may be
more useful.
The ``simplify'' algorithm, which prunes away parts of the graph that 
are not ancestral to a sample, is the key enabling factor for 
this approach to forwards-time simulation and is
described in detail by
\citet{kelleher2018efficient}. It is a 
powerful tool with many other applications,
some of which we discuss in the following sections.

\section*{Converting an eARG to a gARG}
Any eARG can be converted to a gARG without loss of information,
and the conversion process is very similar to
simplifying a prospective gARG with respect to a set of samples.
The process involves two steps, and is illustrated in
Fig.~\ref{fig-ancestry-resolution}.
The first step is straightforward
(Fig.~\ref{fig-ancestry-resolution}B).
We duplicate the nodes and
edges from the input eARG, and the add inheritance annotations
to the output gARG's edges. If the node is a common ancestor node,
we annotate the single outbound edge with the interval $(0,L]$,
for a genome of length $L$.
If the node is a recombination event with a breakpoint $x$,
we annotate the outbound edges with the intervals $(0, x]$
and $(x, L]$.
These inheritance annotations are clearly in one-to-one
correspondance with the information in the input eARG,
but they are a superset of the
``ancestral material''~\citep{wiuf1999ancestry,wiuf1999recombination}
of the sample.

\begin{figure}
\centering
% TODO update the source of this figure to make less tall. It's taking
% up far too much vspace at in native coords at the moment.
\includegraphics[width=0.75\textwidth]{illustrations/ancestry-resolution}
\caption{\label{fig-ancestry-resolution}
Converting the \citet[][Fig.~1]{wiuf1999recombination} example graph
to a gARG. (A) The original eARG; square nodes represent events, with
each recombination (red) showing the corresponding breakpoint.
(B) The corresponding gARG with breakpoints directly converted to
edges annotated with inheritance intervals.
(C) The sample-resolved gARG resulting from simplifying with respect
to the samples, \noderef{a}, \noderef{b}, and \noderef{c}. Dashed lines show edges that are
no longer present (in practice, nodes \noderef{g} and \noderef{q} would also be removed).
Coalescence with respect to the sample is indicated by shaded bars, as
in Fig~\ref{fig-arg-in-pedigree}A.
Line thickness is proportional to the genomic span of each edge.
Nodes representing recombination events have been retained
for clarity, but could be removed by simplification if
desired.
}
\end{figure}

Ancestral material is a critical concept in a retrospective ARG,
and refers to the genomic intervals directly ancestral to the samples
scattered across lineages (the edges of the graph)
at some time in the past. At recombination
events, ancestral material is split between the ancestral lineages.
As genome segments coalesce in a common
ancestor the total amount ancestral material is reduced, until
eventually, as we traverse pastwards through the graph, all
samples reach a most recent common ancestor at every location
along the genome. The process of transmitting ancestral material
along the edges of an ARG is directly analogous to Hudson's
simulation algorithm for the coalescent with
recombination~\citep{hudson1983testing,kelleher2016efficient}.

Fig.~\ref{fig-ancestry-resolution}C shows the graph that we obtain
by resolving the ancestral material transmitted along each edge
using the ``simplify'' algorithm~\citep{kelleher2018efficient}, 
and highlights some important
differences between the eARG and gARG encodings.
Firstly, we can see that some nodes and edges are removed entirely
from the graph by this conversion process. 
The ``grand MRCA'' \noderef{q} is omitted from the
sample resolved gARG because all segments of the genome have
fully coalesced before it is reached. Likewise, the edge
between \noderef{g} and \noderef{j} is omitted because the recombination
event at position $5$ (represented by node \noderef{g})
fell in non-ancestral material.
More generally, we can see that the sample resolved
retrospective gARG of Fig.~\ref{fig-ancestry-resolution}C
allows for ``local'' inspection
of an ARG in a way that is not possible by storing
common ancestor and recombination events in an eARG. Because
the ancestral material is stored with each edge, the
cumulative effects of events over time can be reasoned
about, without first ``replaying'' those events. Many computations
that we wish to perform on an ARG will require resolving
the ancestral material with respect to a sample.
% this is weak, but would be good to emphasise this point.
% you just end up running simplify lots of times, in your
% downstream algs
The gARG encoding
allows us to do this once and to store the result, whereas
the eARG encoding requires us to repeat the process 
each time.

It is important to note here that the \citet{wiuf1999recombination} eARG
in Fig.~\ref{fig-ancestry-resolution} is something of a straw man, 
in that it is unlikely
that an inference or simulation program would produce a retrospective
ARG that contains
nodes or edges that are not ancestral to the sample
(see the discussion of the ``Big ARG'' stochastic process in Appendix XXX,
however). Nonetheless, it
is an instructive example from the literature which highlights several
important properties of ARGs, and the general point about
the need to resolve ancestral material ``on the fly'' for eARG traversals
holds.

@@

\section*{Computational efficiency and data interchange}
ARG inference methods have been successfully applied to vast datasets,
including 500,000 genotyped humans~\citep{kelleher2019inferring,zhang2023biobank}
from UK Biobank~\citep{bycroft2018genome}
and over a million SARS-Cov-2 whole genomes~\citep{zhan2023towards}.
Even larger datasets are
available~\citep[e.g.][]{halldorsson2022sequences} % TODO more
and it seems inevitable that ARGs will be inferred for them.
The efficiency with which we can store and analyse these inferred
ARGs is therefore of great importantance, and a key factor
in the ultimate success of downstream ARG-based methods.
Another key factor in how important ARG-based methods become
across the wide range of predicted application
% TODO more?
areas~\citep{rasmussen2014genome,harris2019database,
hejase2020summary,schaefer2021ancestral,
fan2022genealogical,wohns2022unified,harris2023using,
nowbandegani2023extremely}
is how well inference methods and downstream analysis applications
interoperate. Without a well-defined and efficient shared format
for ARG data interchange,
users will suffer from problems of converting data from one programs
output to another, an unfortunate hallmark of population genetics
software for many years~\citep{excoffier2006computer}.
While current methods tend to tightly couple downstream analysis
of the inferred ARG with the inference
itself within the same software package
% FIXME-  more explicit?
(e.g.\ Relate, ARGneedle)
this approach scales poorly in terms of software development
effort, and is ultimately not compatible with the widespread use
of ARGs for routine data analysis, and a healthy and diverse
software ecosystem.
% FIXME not sure where this goes
Earlier efforts to standardise ARG interchange
did not succeed~\citep{cardona2008extended,mcgill2013graphml}.

The gARG encoding discussed and concretely defined in this manuscript
leads to highly efficient storage and processing of ARG data,
and has already been in use for several years.
The succinct tree sequence data structure (usually known as a ``tree sequence''
for brevity, but see below for confusion around this point)
is a concrete encoding of a gARG, as discussed in this article.
It was originally developed as part of the \texttt{msprime}
simulator~\citep{kelleher2016efficient} and subsequently been
extended and applied to forward-time
simulations~\citep{kelleher2018efficient,haller2018tree},
calculation of population genetics statistics~\citep{ralph2020efficiently}.
The succinct tree sequence encoding extends the basic definition
of a gARG provided here by stipulating a
simple tabular representation and nodes and edges,
and also defining a concise and lossless representation of
sequence variation using the ``site'' and  ``mutation'' tables.
The \texttt{tskit} library~\citep{tskit2023tskit} is a liberally
licensed open source toolkit that provides a comprehensive suite
of tools for working with gARGs (encoded as a succinct tree sequence).
Based on core functionality written
in C, it provides interfaces in C, Python and Rust.
Tskit is mature software, widely used in population genetics, and
has been incorporated into several downstream
applications~\citep[e.g.,][]{haller2019slim,speidel2019method,
adrion2020community,
terasaki2021geonomics,
baumdicker2021efficient,
fan2022genealogical,korfmann2022weak,
mahmoudi2022bayesian,petr2022slendr,rasmussen2022espalier,
zhang2023biobank,nowbandegani2023extremely}.

% FIXME this needs more work and doesn't really fit into the
% narrative right now, but we do need to make this point.
The key insight that makes the succinct tree sequence encoding
an efficient substrate for defining analysis algorithms is that
it allows us to generate the local trees along the genome
in a way that allows us to reason about the \emph{differences}
between those trees.
Sequentially generating the marginal trees along the genome
is also fundamental, and is necessary whenever we need to
perform calculations that are contingent on more than just the
isolated properties of an edge. \cite{kelleher2016efficient}
showed how all trees can be sequentially generated in
constant time per tree transition in a fully simplified gARG.
Furthermore, we can easily reason about how tree topologies
change (and stay the same), leading to efficient algorithms
for computing population genetic
statistics~\citep{kelleher2016efficient,ralph2020efficiently},
implementing the Li and Stephens
model~\citep{kelleher2019inferring,wohns2022unified}
and likely many more.

In retrospect, the choice of terminology around the succinct
tree sequence data structure was unfortunate, and has led to
significant confusion. As first introduced by~\cite{kelleher2016efficient}
a ``tree sequence`` was defined as the set of ``coalescence records''
that is output by Hudson's simulation algorithm.
This was then generalised to use the
tabular encoding described above and formally described as a ``succinct
tree sequence''~\citep{kelleher2018efficient}.
The ``succinct'' prefix here is intended to connect
with the concept of succinct data structures
\citep[e.g.][]{gog2014theory},
which have near-optimal
space usage but still support efficient retrieval and computation.
The concrete data structures of the succinct tree sequence, precisely
defined and with a particular focus on computational efficiency
did not have an obvious connection to the ARG definitions
provided by Griffiths and
colleagues~\citep{griffiths1991two,ethier1990two,
griffiths1996ancestral,griffiths1997ancestral}, and with the lack
of any other precise definitions, it seemed best to avoid the
term to prevent confusion. Unfortunately, however, the opposite
has occured. In describing the output of \tsinfer\ as a ``tree
sequence''~\citep{kelleher2019inferring} many
have---not unreasonably, but incorrectly---concluded that the
output is a collection of \emph{independent}
% MORE?
% https://github.com/tskit-dev/what-is-an-arg-paper/issues/38
trees~\citep[e.g.,][]{hejase2020summary,ignatieva2021kwarg}.
% Could say more here, but maybe we'll have a note about terminology
% in the discussion
